Reproducible Research: Peer Assessment 1
=========================================



#### Loading and preprocessing the data
```{r loadingData, echo=TRUE, results='asis'}

    library(ggplot2)
    library(xtable)

    setwd("~/GitHub/RepData_PeerAssessment1")
    
    rawdata <- read.csv("activity.csv", na.strings=c("NA"))
    
    ## convert dates to date format
    rawdata$date <- as.Date(as.character(rawdata$date))

    ## Section to compute weekday/weekend for one of the problems

    ## declare function to test for weekday or weekend or an input date
    wkEndFcn <- function (inpDt) {
        tmpDay <- weekdays(inpDt) 
        if((tmpDay == "Saturday") || (tmpDay == "Sunday")) 
            return("weekend")
        else
            return("weekday")     
    }

    ### declare column
    DOW <- character(nrow(rawdata))

    ## for loop to do the work
    for (i in 1:nrow(rawdata)) {
        DOW[i] <- wkEndFcn(rawdata$date[i])
    }
 
    # combine data into processed dataset
    procdata <- cbind(rawdata, DOW)

```


### Background Information of activity data:

```{r dataSummary, echo=T, results='asis'}

    xt_sum <- xtable(summary(rawdata))
    print(xt_sum, type="html")


```


The overall mean value for steps is: **`r mean(rawdata$steps, na.rm=T)`**

The overall median value for steps is: **`r median(rawdata$steps, na.rm=T)`**

Notice that there are over **2300** missing values from the data


```{r dataSummaryChart, fig.height=4, fig.width=6,  echo=T}


    day_tot_steps <- aggregate(rawdata$steps, list(day=rawdata$date), FUN=sum, na.rm=TRUE)


    par(mfrow = c(1, 1), mar = c(5, 4, 2, 2))
    barplot(day_tot_steps$x, col="blue", main="Total Steps Each Day", ylab="Step Count", xlab="Days" )


```

**N.B.** From the bar chart above eight days are missing data

______________


### What is mean total number of steps taken per day?
####           Chart of Mean Frequencies for Intevals
```{r , echo=TRUE}
    
##  Create mean step counts for time intervals to create  histogram
    int_mean_steps <- aggregate(rawdata$steps, list(interval=rawdata$interval), FUN=mean, na.rm=TRUE)


    par(mfrow = c(1, 1), mar = c(5, 4, 2, 2))
    hist(int_mean_steps$x, breaks=40, col = "green", xlab="Average Step Counts", ylab="Step Frequencies", main="Histogram of Mean Steps for Intervals")  

```


####          Chart of Median Frequencies for Intervals
```{r , echo=TRUE}

##  Create mean step counts for time intervals to create  histogram
    int_mean_steps <- aggregate(rawdata$steps, list(interval=rawdata$interval), FUN=median, na.rm=TRUE)


    par(mfrow = c(1, 1), mar = c(5, 4, 2, 2))
    hist(int_mean_steps$x, breaks=40, col = "yellow", xlab="Median Step Counts", ylab="Step Frequencies", main="Histogram of Median Steps")  


```


### What is the average daily activity pattern?
```{r, echo=TRUE}

##  Aggregate data for intervals to create frequency histogram
    int_mean_steps <- aggregate(rawdata$steps, list(interval=rawdata$interval), FUN=mean, na.rm=TRUE)

# Find Time Interval for maximum step activity
    max_mean_steps <- int_mean_steps[int_mean_steps$x == max(int_mean_steps$x),]


    par(mfrow = c(1, 1), mar = c(5, 4, 2, 2))
    plot(int_mean_steps$interval, int_mean_steps$x, type="l", col = "blue", xlab="Time Intervals", ylab="Average Steps Taken", main="Time Series of Avg. Steps at Each Time Interval")  


```

#### The maximum average steps is: **`r round(max_mean_steps$x)`** which occured at interval:  **`r max_mean_steps$interval`**

### Imputing missing values
```{r, echo=TRUE}

imputeVal <- function(inpSteps) {
    replace(inpSteps, is.na(inpSteps), mean(rawdata$steps, na.rm = TRUE))
}

```


### Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}


```